
# Data structure{.tabset}
We'll use a fake demo data set containing the number of phone calls between the president of some EU countries.


```{r}
counties <- rKenyaCensus::CountyGPS

mpesa <- tibble(
  source=sample(counties$CountyCode, size= 3*nrow(counties), replace = T),
  destination= sample(counties$CountyCode, size= 3*nrow(counties), replace = T),
  amount=sample(2:14, size= 3*nrow(counties), replace = T)
)
head(mpesa)
```


To visualize the network graph, we need to create two data frames:
 
    * nodes list: containing nodes labels and other nodes attributes
    * edges list: containing the relationship between the nodes. It consists of the edge list and any additional edge attributes.


## Create nodes list
You  need to check for any data manipulation b4 proceeding.

```{r}

source <- mpesa %>% 
  distinct(source) %>% 
  rename(county=source) #get distinct source countries.

destination <- mpesa %>% 
  distinct(destination) %>% 
  rename(county=destination) #get distinct destination countries.

nodes <- full_join(source, destination, by="county")# creating nodes dataset.

nodes <- nodes %>% 
  rowid_to_column("id")
```

##  Create edges list
Key steps:
1. Take the phone.call data, which are already in edges list format, showing the connection between nodes. Rename the column "n.call" to "weight".
2. Join the node IDs to the edges list data
a. Do this for the "source" column and rename the id column that are brought over from nodes. New name: "from".
b. Do this for the "destination" column and rename the id column. New name: "to"
c. Select only the columns "from" and "to" in the edge data. We don't need to keep the column "source" and "destination" containing the names of countries. These information are already present in the node
data.


```{r}
transaction <- mpesa %>% 
  rename(weight=amount)# renaming the nc.call variable

edges <- transaction%>% 
  left_join(nodes, by=c("source"="county")) %>% 
  rename(from=id)

  
edges <- edges %>% 
  left_join(nodes , by=c("destination"="county")) %>% 
  rename(to=id)

edges <- select(edges, from, to, weight)

head(edges, 3)
```
