
# DATA STRUCTURE{.tabset}
We'll use an hypothetical data of mpesa transaction i created, now data thats 0one of the major forms of money tranfer in Kenya and East Africa by [Safaricom PLC]("https://www.safaricom.com")



```{r mpesa}
mpesa <- read_csv("mpesa.csv") #reading in the data

```
## Data anatomy.
sometimes we may want to check how data looks and perform any necessary data manipulation.
check ......for description of the data.
```{r}
head(mpesa)
str(mpesa)
```

To visualize the network graph, we need to create two data frames:
 
    * nodes list: containing nodes labels and other nodes attributes
    * edges list: containing the relationship between the nodes. It consists of the edge list and any additional edge attributes.


## Create nodes list

Lets create nodes from unique counties in the data.
```{r nodes}

source <- mpesa %>% 
  distinct(source) %>% 
  rename(county=source) #get distinct source counties.

destination <- mpesa %>% 
  distinct(destination) %>% 
  rename(county=destination) #get distinct destination counties.

nodes <- full_join(source, destination, by="county")# creating nodes dataset.

nodes <- nodes %>% 
  rowid_to_column("id")


```

##  Create edges list
Key steps:
1. Take the mpesa data, which are already in edges list format, showing the connection between nodes. Rename the column "amount" to "weight".
2. Join the node IDs to the edges list data

     a. Do this for the "source" column and rename the id column that are brought over from nodes. New name: "from".
     b. Do this for the "destination" column and rename the id column. New name: "to"
    c. Select only the columns "from" and "to" in the edge data. We don't need to keep the column "source" and "destination" containing the names of countries. These information are already present in the node
data.


```{r edges}
transaction <- mpesa %>% 
  rename(weight=amount)# renaming the nc.call variable

edges <- transaction%>% 
  left_join(nodes, by=c("source"="county")) %>% 
  rename(from=id)

  
edges <- edges %>% 
  left_join(nodes , by=c("destination"="county")) %>% 
  rename(to=id)

edges <- select(edges, from, to, weight)

head(edges, 3)
```
