# TOOLS OF VISUALIZATION {.tabset .tabset-fade .tabset-pills #vis}

## __igraph__

```{r}
net.igraph <- graph_from_data_frame(
d = edges, vertices = nodes,
directed = TRUE
)
net.igraph

plot(net.igraph)
```
The description of our igraph object ( _net.igraph_ )starts with three letters:

    *D or U, for a directed or undirected graph
    *N for a named graph (where nodes have a name attribute)
    *W for a weighted graph (where edges have a weight attribute)
    
The two numbers that follow ($47$, $141$) refer to the number of nodes and edges in the graph. The description also lists node & edge attributes, for example:

(v/n) - vertex-level numeric attribute
(v/c) - vertex-level character attribute
(e/n) - edge-level numeric attribute

This concept makes it possible hence to access the various attributes of the graph making it possible manipulation.
```{r graph_attributes}
E(net.igraph)       # The edges of the "net.igraph" object
V(net.igraph)       # The vertices of the "net.igraph" object


```
When we plotted net.igraph above we noticed the graph was a mess, lets try make it presentable
First we remove loops.
secondly we can group those alike edges.
```{r}
net.igraph_clean <- simplify(net.igraph, remove.multiple = FALSE, remove.loops = TRUE, edge.attr.comb = list(Weight="sum", "ignore")) 

plot(net.igraph_clean, 
     layout = layout_with_graphopt,
     edge.arrow.size = 0.2,
     edge.color="blue",
     vertex.color="green",
     vertex.frame.color="#ffffff"
     )
```




##  __Tidygraph & ggraph__

_***tidygraph***_ and _***ggraph***_ represent an attempt to bring network analysis into the tidyverse workflow. For this reason tidygraph objects are masks of igraph objects.
However using the two, you sacrifice few of the network analysis capabilities of igraph.
When we print a _tbl_graph_ we see this __activate__ message by default on node data. we can change to edge data using __activate()__ for us to manipulate the data like rearranging etc.


__ggraph__ have special geoms as compared to ggplot; _geom_node_ and _geom_edge_. 
The ggraph theme can be set for a series of plots with the _set_graph_style()_ command run before the graphs are plotted or by using _theme_graph()_ in the individual plots. Here, I will use the latter method.

```{r tidygraph}

# creating a tbl_graph

net_tbl <- tbl_graph(nodes=nodes, edges = edges, directed = T)

# Or creating from igraph

######net_tbl2 <- as_tbl_graph(net.igraph)

# Printing
net_tbl

# activate the edge list

net_tbl %>% 
  activate(edges) %>% 
  arrange(desc(weight))
```



```{r ggraph}
# Plotting using ggraph
net_tbl %>% 
  ggraph(layout ="graphopt" )+
  geom_node_point()+
  geom_edge_link(aes(width=weight), alpha=0.8)+ 
  scale_edge_width(range = c(1, 2))+
  geom_node_text(aes(label=county), repel = T)+
  labs(edge_width = "transaction") +
  theme_graph()
```


##   __Interactive network graphs with visNetwork and networkD3__

The htmlwidgets set of packages makes it possible to use R to create interactive JavaScript visualizations. The disadvantage of working with both visNetwork and networkD3 is that they expect edge lists and node lists to use specific nomenclature.
 When creating an interactive plot with __visnetwork()__ you must have a node list with _'ID'_ column and an edge list with _'from' _ and _'to'_ columns, the function also plots the labels for the nodes, using the names of the cities from the “label” column in the node list.
 
Visnetwork can also be build from an igraph object using _visIgraph()_ function. visNetwork() uses column names from the edge and node lists to plot network attributes instead of arguments within the function call.  This means that it is necessary to do some data manipulation to get a “width” column in the edge list. The width attribute for visNetwork() does not scale the values, so we have to do this manually. Both of these actions can be done with the mutate() function and some simple arithmetic.
 

To make a networkD3 graph with a edge and node list requires that the IDs be a series of numeric integers that begin with 0.
The forceNetwork() function uses a series of arguments to adjust the graph and plot network attributes. The “Links” and “Nodes” arguments provide the data for the plot in the form of edge and node lists. The function also requires “NodeID” and “Group” arguments.


```{r}

edges <- mutate(edges, width = weight/3 + 1)

vis_net <- visNetwork(nodes = nodes, edges = edges)

vis_net %>% 
  visIgraphLayout(layout = "layout_with_fr") %>% 
  visEdges(arrows = "to",
           shadow = T)
```


```{r}
#subracting 1 to force the id's to begin at 0.

nodes_d3 <- mutate(nodes, id = id - 1)
edges_d3 <- mutate(edges, from = from - 1, to = to - 1)

# groups help to cluster similar values to something like a family.

forceNetwork(Links = edges_d3, Nodes = nodes_d3, Source = "from", Target = "to", 
             NodeID = "county", Group = "id", Value = "weight", 
             opacity = 1, fontSize = 16, zoom = TRUE)
```

## Communities
we go further and try find aspects such as the mean and standard deviationof a dense link like the one we are having. which will further help us find communities

```{r}
hist(edges$weight)
mean(edges$weight)
sd(edges$weight)

#lets create a communities based on mean of the graph.
cut.off <- mean(edges$weight) 
net.mean <- delete_edges(net.igraph_clean,
                         E(net.igraph_clean)[weight<cut.off])

plot(net.mean, 
     layout=layout_with_kk, 
     edge.arrow.size=0.2) 


# Community detection:
comm <- cluster_edge_betweenness(net.igraph_clean)
class(comm)

V(net.igraph_clean)$community <- comm$membership
colrs <- adjustcolor( c("gray50", "tomato", "gold", "yellowgreen"), alpha=.6)
plot(net.igraph_clean, vertex.color=colrs[V(net.igraph_clean)$community])


```



















