---
title: "NETWORK SCIENCE"
author: DIR-TERRANCE^[FOUNDER:RWILLS STATISTICAL COMPANY,EMAIL:consultancyrwillsstats@gmail.com]
date: "`r Sys.Date()`"
site: bookdown::bookdown_site
documentclass: book
bibliography: [book.bib, packages.bib]
biblio-style: apalike
link-citations: yes
description: "This is an introduction tutorial of Network Science written in R."
---


# Prerequisites {.unnumbered}

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, warning = F, message = F)
```

## Packages.

Install the following packages to follow the book easily.

```{r eval=FALSE}
if(!require(devtools))(install.packages("devtools"))
devtools::install_github("kassambara/navdata")
install.packages("bookdown")
install.packages("tidyverse")
install.packages("igraph") 
install.packages("network") 
install.packages("sna")
install.packages("ggraph")
install.packages("visNetwork")
install.packages("threejs")
install.packages("networkD3")
install.packages("ndtv")
```


```{r eval=T}
library(bookdown)
# 
library(tidyverse)
#

library(rticles)
#
library("igraph") 
library("network") 
library("sna")
library("ggraph")
library("visNetwork")
library("threejs")
library("networkD3")
library(rKenyaCensus)
#library("ndtv")

library(tidygraph)

```

### Bibliography

automatically create a bib database for R packages

```{r include=T}
knitr::write_bib(c(
  .packages(), 'bookdown', 'knitr', 'rmarkdown', 'tidyverse','rKenyaCensus'
), 'packages.bib')
```

# **INTRODUCTION**
Network Analysis is used to investigate and visualize the inter-relationship
between entities (individuals, things).

Network analysis extensively uses graph theory.

## **Graph theory**: Network analysis
Network graphs are mainly made of Nodes and edges.

 * *nodes*: The entities (individual actors, people, or things) to be connected in the network. Synonyms: vertices of a graph.
 
 * *edges*: The connections (interactions or relationships) between the entities.Synonyms: links, ties.
 
 * *adjacency matrix*: a square matrix in which the column and row names are the nodes of the network. This is a standard data format accepted by many network analysis packages in R. Synonyms: *sociomatrices*. Within the matrix a 1 specifies that there is a link between the nodes, and a 0 indicates no link.
 
 - *edge list*: a data frame containing at least two columns: one column of nodes corresponding to the source of a connection and another column of nodes that contains the target of the connection. The nodes in the data are identified by unique IDs.
 
- *Node list*: a data frame with a single column listing the node IDs found in the edge list. You can also add attribute columns to the data frame such as the names of the nodes or grouping variables.

- *Weighted network graph*: An edge list can also contain additional columns describing attributes of the edges such as a magnitude aspect for an edge. If the edges have a magnitude attribute the graph is considered weighted.

- *Directed* and *undirected network graph*: 

    + If the distinction between source and target is meaningful, the network is directed. Directed edges represent an ordering of nodes, like a relationship extending from one nodes to another, where switching the direction would change the structure of the network. The World Wide Web is an example of a directed network because hyperlinks connect one Web page to another, but not necessarily the other way around (Tyner, Briatte, and Hofmann 2017).

    + If the distinction is not meaningful, the network is undirected. Undirected edges are simply links between nodes where order does not matter. Coauthorship networks represent examples of undirected networks, where nodes are authors and they are connected by an edge if they have written a publication together (Tyner, Briatte, and Hofmann 2017).



































































































































































































































































































